---
title: "Your Data Storytelling Project Title"
subtitle: "DANL 310: Data Visualization and Presentation"
author:
  - name: Aaron Zalen
    affiliation: SUNY Geneseo
date: last-modified

execute:
  echo: true
  eval: true
  warning: false
  message: false
  fig-width: 9
  fig-height: 6

format:
  html:
    toc: true
    toc-title: "Contents"
    toc-depth: 4
    toc-expand: 3
    number-sections: false
    code-fold: false
    code-tools: true
    highlight-style: atom-one
---

```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(hrbrthemes)
library(rmarkdown)
library(dplyr)
library(ggplot2)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggthemes)
library(maps)

# Add other packages as needed.
# Examples:
# library(lubridate)
# library(janitor)
# library(scales)
# library(viridis)
# library(plotly)
# library(DT)
# library(sf)

# Feel free to revise this theme.
theme_set(
  theme_ipsum() +
    theme(
      strip.background = element_rect(fill = "lightgray"),
      axis.title.x = element_text(size = rel(1.33)),
      axis.title.y = element_text(size = rel(1.33))
    )
)

# Optional color-blind-friendly scales.
# scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
# scale_fill_discrete <- function(...) scale_fill_viridis_d(...)
```

# This page is currently a work in progress

# Project Overview

- **Project Webpage:** [HSR Visualization project](https://1aaronz.github.io/danl-310-data-storytelling.html)

This project visualizes Global High Speed Rail (HSR) progress compared to the US.

Your project should tell an interesting and coherent story using:

- descriptive statistics,
- data transformation, and
- data visualization.

A literature review, regression model, machine learning model, or formal statistical model is **not required**.

# Big Idea

Despite public interest in HSR in the US, 60% of U.S. voters have a favorable opinion of high-speed rail Only 16% view it unfavorably (Public Policy Polling), the US has fallen behind in HSR construction when compared to countries that have built HSR.

This project emphasizes the difference in the length of HSR other countries have managed to achieve while the US continues to have slow progress in HSR construction. This project will compare both historical HSR lengths and visualize the change up to the present day and 10 years into the future.

# Guiding Question or Story Theme

The guiding question of this project is: "How has the length of HSR lines by country changed overtime?"

# Background and Context

What qualifies as "high speed rail" can be ambiguous. For the purposes of this project, high speed rail is defined as "the length of track where passenger trains can reliably go at least 200 kilometers per hour". In the case of the United States, the North East Corridor line can in theory carry high speed rail much further than what the dataset for this project provides, however only about 80km of the North East Corridor is currently operating high speed rail, so only 80km is counted for this project.

# Data

```{r data-import}
hsr <- read.csv("hsr.csv") %>% 
  select(Line, Length_km, Opening_year, country, Status, Maximum.speed)

paged_table(hsr)
```

## Key Variables

Line: Name of high speed rail line

Length_KM: Length of the high speed rail line in Kilometers

country: Name of the country the rail line belongs to

Opening_year: The year the high speed rail on the line became operational

## Unit of Observation

Each row represents one rail line capable of carrying high-speed rail.

## Data Source

The data was scraped from tables from wikipedia on 4/30/2026

# Data Preparation and Transformation

Describe the main steps you used to prepare the data for storytelling.

This may include:

- cleaning missing or invalid values,
- renaming variables,
- filtering observations,
- creating new variables,
- grouping and summarizing data,
- reshaping data from wide to long format, or long to wide format,
- joining multiple datasets, and
- preparing data specifically for your Shiny app or dashboard.

```{r data-preparation}
# Replace this example with your own data cleaning and transformation code.

hsr_all <- read.csv("hsr_all.csv")
# ------------------------------------------------------------
# 5. CLEAN LENGTH COLUMN
# ------------------------------------------------------------
if ("Length" %in% names(hsr_all)) {
  hsr_all <- hsr_all %>%
    mutate(
      Length_clean = gsub(",", "", Length),
      Length_km = as.numeric(str_extract(Length_clean, "[0-9.]+"))
    )
}

# ------------------------------------------------------------
# 6. CLEAN OPENING YEAR (ROBUST FIX)
# ------------------------------------------------------------
if ("Opening" %in% names(hsr_all)) {
  hsr_all <- hsr_all %>%
    mutate(
      Opening_year = as.numeric(str_extract(Opening, "\\d{4}"))
    )
}

hsr <- hsr_all %>% 
  select(Line, Length_km, Opening_year, country, Status, Maximum.speed)
```

The the length and opening year columns had to have the characters removed from them and set as numeric columns.

```{r}
hsr_2026 <- hsr %>% 
  filter(Opening_year<=2026) %>% 
  group_by(country) %>% 
  summarise(total_km = sum(Length_km))


hsr_2000 <- hsr %>% 
  filter(Opening_year<=2000) %>% 
  group_by(country) %>% 
  summarise(total_km = sum(Length_km))

hsr_2000 <- hsr_2000 %>%
  mutate(
    country = as.character(country),
    is_us = country == "United_States")
```

additional cleaning for certain years and creating a flag for the US to highlight on the maps.

# Descriptive Statistics

Use descriptive statistics to help your audience understand the data before you tell the main story.

## Basic Summary

```{r summary-statistics}
skim(hsr)
```

## Important Counts, Rates, or Group Summaries

```{r descriptive-summary}
# Example structure:

hsr |>
  group_by(country) |>
  filter(Opening_year<=2026) %>% 
  summarize(
    n = n(),
    mean_HSR_line = mean(Length_km, na.rm = TRUE),
    Total_HSR_km = sum(Length_km, na.rm = TRUE)
  ) |>
  arrange(desc(Total_HSR_km))
```

## Initial Observations

Briefly explain what the descriptive statistics show.

Focus on patterns such as:

- largest or smallest groups,
- typical values,
- variation across groups,
- outliers or unusual cases,
- missing data issues, and
- patterns that set up your main story.

China appears as a big outlier in total HSR kilometers. The US is only seen much further down the list, almost all countries that have any HSR are ahead of the US.

# Data Visualization and Storytelling

This section is the main body of your project. Use visualizations to build a clear story.

Each visualization should include:

- a clear, message-driven title,
- readable axis labels,
- appropriate chart type,
- thoughtful color choices,
- colorblind-friendly colors,
- a source note or caption when appropriate, and
- a short written interpretation.

Your webpage should feel like one coherent narrative, not a disconnected collection of plots.

## Visualization 1: HSR by country in 2000

```{r visualization-1}
# Replace this example with your own visualization.

hsr_2000 <- hsr_2000 %>%
  mutate(
    country = fct_reorder(country, total_km)
  )

ggplot(hsr_2000, aes(
  x = total_km,
  y = fct_reorder(country, total_km)
)) +
  geom_col(aes(fill = country)) +
  scale_y_discrete(labels = function(x) {
    ifelse(x == "United_States", "United_States", x)
  }) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.y = element_text(
      color = ifelse(levels(fct_reorder(hsr_2000$country, hsr_2000$total_km)) == "United_States",
                     "red", "black"),
      face = ifelse(levels(fct_reorder(hsr_2000$country, hsr_2000$total_km)) == "United_States",
                    "bold", "plain")
    )
  ) +
  labs(
    title = "Total Operational High Speed Rail Per Country in 2000",
    y = "",
    x= "Total HSR in Km",
    subtitle = "Countries Ahead of the US: 13"
  )+
  scale_fill_viridis_d()
```

**Interpretation:**\
Explain what this visualization shows. What should the audience notice first? How does this figure begin the story?

## Visualization 2: HSR by Country in 2026

```{r visualization-2}
hsr_2026 <- hsr_2026 %>%
  left_join(
    hsr_2000 %>% select(country, total_km) %>% rename(km_2000 = total_km),
    by = "country"
  ) %>%
  mutate(
    km_2000 = ifelse(is.na(km_2000), 0, km_2000),
    new_hsr = km_2000 == 0
  )

hsr_2026 <- hsr_2026 %>%
  mutate(
    country = fct_reorder(country, total_km)
  )

ggplot(hsr_2026, aes(
  x = total_km,
  y = fct_reorder(country, total_km)
)) +
  geom_col(aes(fill = country)) +
  scale_y_discrete(labels = function(x) {
    x
  }) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.y = element_text(
      color = ifelse(
        levels(fct_reorder(hsr_2026$country, hsr_2026$total_km)) == "United_States",
        "red",
        "black"
      ),
      face = ifelse(
        levels(fct_reorder(hsr_2026$country, hsr_2026$total_km)) %in% 
          hsr_2026$country[hsr_2026$new_hsr],
        "bold",
        "plain"
      )
    )
  ) +
  labs(
    title = "Total Operational High Speed Rail Per Country in 2026",
    y = "",
    x = "Total HSR in Km",
    subtitle = "Countries with no HSR in 2000 are Bolded | Countries Ahead of the US: 26"
  ) +
  scale_fill_viridis_d()
```

**Interpretation:**\
Explain the key comparison. What is similar, different, surprising, or important?

## Visualization 3: Future Prospects, the 10 year projection of new HSR  by country

```{r visualization-3}
hsr_clean <- hsr_2026 %>%
  mutate(
    country = dplyr::recode(country,
                     "United_States" = "United States of America",
                     "United_Kingdom" = "United Kingdom",
                     "South_Korea" = "South Korea",
                     "Saudi_Arabia" = "Saudi Arabia",
                     "Russia" = "Russia"
    )
  )

world <- ne_countries(scale = "medium", returnclass = "sf")

us_value <- hsr_clean %>%
  filter(country == "United States of America") %>%
  pull(total_km)

world_hsr <- world %>%
  left_join(hsr_clean, by = c("name" = "country")) %>%
  mutate(
    status = case_when(
      name == "United States of America" ~ "United States",
      total_km > us_value ~ "Above US",
      TRUE ~ "Below US"   # <- THIS is the key fix
    )
  )

ggplot(world_hsr) +
  geom_sf(aes(fill = status), color = "white", linewidth = 0.2) +
  scale_fill_manual(
    values = c(
      "Above US" = "#D55E00",
      "Below US" = "#999999",
      "United States" = "#0072B2"
    )
  ) +
  labs(
    title = "Countries With More Operational High Speed Rail than the US (2026)",
    fill = ""
  )+
  theme_map()+
  theme(title = element_text(size=15),
        axis.text = element_blank(),
        legend.position = c(0, 0.5), 
        legend.justification = c(0, 0.5),
        legend.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5))
```

**Interpretation:**\
Explain how this visualization deepens the story. How does it connect to your Big Idea?

## Optional Visualization 4: Additional Evidence 1

```{r visualization-4}
# Optional additional figure.
```

**Interpretation:**\
Use this only if it adds something meaningful to the story.

## Optional Visualization 5: Additional Evidence 2

```{r visualization-5}
# Optional additional figure.
```

**Interpretation:**\
Use this only if it adds something meaningful to the story.

## Shiny App

```{=html}
<div style="display: flex; justify-content: center; margin: 1em 0;">
  <iframe
    src="YOUR_SHINY_APP_URL"
    width="800"
    height="800"
    frameborder="0"
    style="border: none;"
  ></iframe>
</div>
```

Briefly describe the Shiny app and dashboard that accompany this project.

## Dashboard

- **Dashboard:** [Insert link here](YOUR_DASHBOARD_URL)

Explain how dashboard help users interact with or explore the project data.

# Connecting the Dots

Bring the visual evidence together.

Explain how your descriptive statistics, transformations, and visualizations work together to answer your guiding question or support your story theme.

You may discuss:

- the strongest pattern in the data,
- a surprising result,
- differences across groups,
- changes over time,
- geographic or categorical variation,
- practical, business, economic, social, cultural, or policy relevance,
- why your audience should care.

# Discussion

Interpret the meaning of your findings in plain language.

This project does **not** require regression, machine learning, or causal inference. Avoid claiming that one variable causes another unless your project design truly supports that claim.

Use careful language such as:

- The data suggest that...
- The pattern is consistent with...
- One possible explanation is...
- This comparison shows that...
- This trend raises questions about...

## Limitations

Discuss what your data story can and cannot show.

Possible limitations include:

- missing data,
- limited time period,
- measurement issues,
- small sample size,
- data not representing all populations,
- descriptive analysis cannot prove causality.

# Conclusion

Summarize your project in one or two paragraphs.

Return to your Big Idea and explain the main takeaway your audience should remember.

A strong conclusion should answer:

- What did the data reveal?
- Why does it matter?
- What should the audience remember after leaving your webpage?

# Future Directions

Briefly explain what you would explore next.

Examples:

- additional variables,
- a longer time period,
- different geographic areas,
- more detailed group comparisons,
- a dashboard or interactive visualization improvement,
- a follow-up analysis using modeling, if appropriate.

# References

Public Policy Polling. (2024). National survey on high-speed rail (conducted for the U.S. High Speed Rail Coalition). https://www.ushsr.com/

Wikipedia contributors. (n.d.). List of high-speed railway lines. In Wikipedia. Retrieved May 6, 2026, from https://en.wikipedia.org/wiki/List_of_high-speed_railway_lines

# Appendix

Use the appendix for extra materials that are useful but not central to the story.

Examples:

- additional tables,
- additional figures,
- extra data cleaning details,
- code that is too long for the main narrative.

Code to scrape the data

```{r appendix-code}
library(rvest)
library(dplyr)
library(purrr)
library(stringr)
library(xml2)
# ------------------------------------------------------------
# 1. SCRAPE PAGE
# ------------------------------------------------------------
url <- "https://en.wikipedia.org/wiki/High-speed_rail_by_country"
page <- read_html(url)

nodes <- xml_find_all(page, "//h2 | //h3 | //table")

tables <- list()
current_name <- NULL
i <- 1

for (node in nodes) {
  
  # update section name
  if (xml_name(node) %in% c("h2", "h3")) {
    current_name <- xml_text(node, trim = TRUE)
  }
  
  # store table under current section name
  if (xml_name(node) == "table") {
    df <- html_table(node, fill = TRUE)
    
    df <- as.data.frame(df, stringsAsFactors = FALSE)
    df[] <- lapply(df, as.character)  # 🔥 FIX TYPE CONSISTENCY
    
    tables[[i]] <- df
    names(tables)[i] <- current_name
    i <- i + 1
  }
}

# ------------------------------------------------------------
# 3. MANUAL COUNTRY ASSIGNMENT
# ------------------------------------------------------------
egypt <- tables[[5]]
morocco <- tables[[6]]
china <- tables[[7]]
india <- tables[[8]]
indonesia <- tables[[9]]
japan <- tables[[10]]
saudi_arabia <- tables[[11]]
south_korea <- tables[[12]]
taiwan <- tables[[13]]
thailand <- tables[[14]]
turkey <- tables[[15]]
uzbekistan <- tables[[16]]
austria <- tables[[17]]
baltic_states <- tables[[18]]
belgium <- tables[[19]]
czech_republic <- tables[[20]]
denmark <- tables[[21]]
finland <- tables[[22]]
france <- tables[[23]]
germany <- tables[[24]]
greece <- tables[[25]]
italy <- tables[[26]]
netherlands <- tables[[27]]
norway <- tables[[28]]
poland <- tables[[29]]
portugal <- tables[[30]]
russia <- tables[[31]]
serbia <- tables[[32]]
spain <- tables[[33]]
sweden <- tables[[34]]
switzerland <- tables[[35]]
united_kingdom <- tables[[36]]
united_states <- tables[[37]]

hsr_list <- list(
  Egypt = egypt,
  Morocco = morocco,
  China = china,
  India = india,
  Indonesia = indonesia,
  Japan = japan,
  Saudi_Arabia = saudi_arabia,
  South_Korea = south_korea,
  Taiwan = taiwan,
  Thailand = thailand,
  Turkey = turkey,
  Uzbekistan = uzbekistan,
  Austria = austria,
  Baltic_States = baltic_states,
  Belgium = belgium,
  Czech_Republic = czech_republic,
  Denmark = denmark,
  Finland = finland,
  France = france,
  Germany = germany,
  Greece = greece,
  Italy = italy,
  Netherlands = netherlands,
  Norway = norway,
  Poland = poland,
  Portugal = portugal,
  Russia = russia,
  Serbia = serbia,
  Spain = spain,
  Sweden = sweden,
  Switzerland = switzerland,
  United_Kingdom = united_kingdom,
  United_States = united_states
)

# ------------------------------------------------------------
# 4. COMBINE ALL TABLES (SAFE)
# ------------------------------------------------------------
hsr_all <- bind_rows(
  lapply(names(hsr_list), function(country) {
    df <- hsr_list[[country]]
    df$country <- country
    df
  }),
  .id = "table_id"
)

# write.csv(hsr_all, "hsr_all.csv")
```

# Project Checklist

Before submitting, check that your project includes the following:

- [ ] A clear Big Idea or main message.
- [ ] A clear guiding question or story theme.
- [ ] A description of the data source and unit of observation.
- [ ] Data transformation code and explanation.
- [ ] Descriptive statistics that help set up the story.
- [ ] Clear and readable visualizations.
- [ ] Written interpretation after each major figure.
- [ ] A coherent narrative connecting the figures together.
- [ ] A conclusion explaining why the story matters.
- [ ] A published project webpage on each member's personal GitHub website.
- [ ] A published Shiny app using the project data.
- [ ] A published dashboard using the project data.
- [ ] No unnecessary regression, machine learning, or formal modeling section.
- [ ] No major code errors, warnings, or redundant output.
