---
title: DANL Project
subtitle: "Data-Driven Mastery: Unlocking Business Potential"
author: 
  - Name 1
  - Name 2
  - Name 3
  - Name 4
  - Name 5

toc: true
toc-title: "Table of Contents"
toc-depth: 2
number-sections: true

fig-width: 9

execute:
  code-tools: true
  message: false
  warning: false

from: markdown+emoji
---

```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```

# Introduction

The goal intended of the research and analysis of the stocks performed was to help people understand what certain stocks would classify as, whether it be a good investment overall, something preferable for the short/long run, what the volatility of a stock is, and other performance functions that would help the reader grasp what to look for when investing.

```{r}

```

# Data

We retrieved the dataset \"Stock Market: Historical Data of Top 10 Companies\" from Kaggle. <https://www.kaggle.com/datasets/khushipitroda/stock-market-historical-data-of-top-10-companies>

The data includes 7 variables and 25160 observations. The variables are Company which is a factor variable for each company name. As well as variables that show the price at a given time.

## Summary Statistics

```{r}
Stock_data_clean <- Stock_data %>% 
  mutate(Close.Last.Numeric = gsub("\\$","", Stock_data$Close.Last),
         Open.Numeric = gsub("\\$","", Stock_data$Open),
         High.Numeric = gsub("\\$","", Stock_data$High),
         Low.Numeric = gsub("\\$","", Stock_data$Low)) %>% 
  mutate(Close_Last = as.numeric(Close.Last.Numeric),
         Open_Numeric = as.numeric(Open.Numeric),
         High_Numeric = as.numeric(High.Numeric),
         Low_Numeric  = as.numeric(Low.Numeric)) %>% 
  select(Company, Date, Volume, Close_Last:Low_Numeric)

```

```{r}
#| results: asis
#| echo: false
rmarkdown::paged_table(mpg) 
```

```{r}
skim(Stock_data_clean) %>% 
  select(-n_missing)
```

## Research Question 1: Which Stocks Have the Most Historical Volatility

The following boxplot shows how the distribution of highway MPG (`hwy`) varies by a type of cars (`class`) :blue_car: :truck: :minibus:.

```{r}
ggplot(data = mpg) +
  geom_boxplot(aes(x = class, y = hwy, fill = class),
               show.legend = F) +
  labs(x = "Class", y = "Highway\nMPG") 
```
