---
title: Data Analytics Capstone Project
subtitle: "NCAA March Madness Prediction"
author:
  - name: Aaron Zalen 
    affiliation: SUNY Geneseo
    
date: last-modified

execute: 
  echo: true
  eval: true
  warning: false
  message: false
  fig-width: 10
  fig-height: 6
  dir: project

format:
  html:
    code-fold: true
    toc: true
    toc-title: "Contents"
    toc-expand: 3
    number-sections: true
    code-tools: true
    highlight-style: atom-one
---

```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(rmarkdown)
library(DT)
library(dplyr)
library(readr)
library(purrr)
library(tibble)
library(ranger)  
library(stringr)
library(tidyr)
library(glmnet)
library(dplyr)
library(stringdist)
library(readr)
library(purrr)
library(tibble)
# library(randomForest)
library(stringr)
library(tidyr)
library(glmnet)
library(tidyverse)

# Feel free to revise this ggplot theme_set
theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = ggplot2::margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 90,
                               size = rel(1.33),
                               margin = ggplot2::margin(0,10,0,0))
                )
          )


# Feel free to use below global options for color-blind-friendly scales
# scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
# scale_fill_discrete <- function(...) scale_fill_viridis_d(...)
```

## This page is currently under construction

# Introduction

March Madness is an annual tournament for division 1 college basketball. I will be training prediction models to assign a probability to teams winning in a match up. One of the prediction models is derived from human-selected bracket data sourced from ESPN to compare the results of the machine learning models against.

## Research Question

The research question this project aims to answer is "How do NCAA March Madness prediction models perform relative to human-selected brackets?"

To answer this question, I will be evaluating each model on Brier-score, Log-loss, and Scored Simulated Brackets

The goal is to find a way to fill out brackets which consistently outperforms both people and the other prediction models. If a model is able to consistently outperform people at filling out brackets, anyone who relies on that model will have an advantage.

# Background and Related Literature

The structure of March Madness is a 64 team, single elimination tournament.

Millions of people fill out brackets each year, attempting to predict which teams will move on to each round. Billions of dollars each year are gambled on the tournament.

# Data

## Data Source

Regular season performance data and bracket structure data: Kaggle

Advanced efficiency metrics: Kenpom

Public picks data: mRchmadness package, which scrapes data from ESPN

## Loading the data

There is a large number of csv files that could be useful for the analysis, so I load the data into a directory and pull each

```{r}
data_dir <- "csvs"

files <- list.files(
  data_dir,
  pattern = "\\.csv$",
  full.names = TRUE
)

data <- files %>%
  set_names(tools::file_path_sans_ext(basename(.))) %>%
  map(~ tryCatch(
    read_csv(.x, show_col_types = FALSE),
    error = function(e) {
      message("Error reading: ", .x)
      NULL
    }
  ))

kenpom_final <- data$kenpom_tournament_ready

data
```

## Data summary

### Regular season stats

```{r}
skim(data$MRegularSeasonDetailedResults)
```

### Advanced efficiency metrics

```{r}
skim(data$INT_KenPom_Summary_Pre_Tourny)
```

### Public pick data

```{r}
skim(data$"Public Picks")
```

## Unit of Observation

Each observation of MRegularSeasonDetailedResults represents game-level data for each match-up in the regular season.

Each observation of INT_KenPom_Summary_Pre_Tourny represents team-level pre-tournament aggregate data

## Key Variables

The outcome variable we are predicting for is win probability this following are the variables used to build difference features for each machine learning model:

WinPct= Win percentage

AvgPoints= Average points

AvgOppPoints= Average Opponent Points

FG2_Pct= two-point percentage

FG3_Pct= three-point percentage

FT_Pct= free-throw percentage

OR_pg= Offensive rebounds per game

DR_pg= Defensive rebounds per game

TO_pg= Turnovers per game

CareerTourneyWins= Coach's historical success in previous tournaments, measured in number of NCAA tournament games won.

AdjEM= Adjusted efficiency margin, the number of points a team is expected to outscore the average D-1 team over 100 possessions

SOV= Strength of victory, average AdjEM percentile of opponents that a team defeated

AdjOE= Adjusted offensive efficiency, sourced from kenpom, points scored per 100 possessions and adjusted for opponent.

AdjDE= Adjusted defensive efficiency, points allowed per 100 possessions and adjusted for opponent.

RankAdjEM= Ranked adjusted efficiency margin, the number ranking of each team's Adjusted efficiency margin each season

AdjTempo= Adjusted tempo, measured in possessions per 40 minutes and adjusted for opponent.

## Data Cleaning

### Creating the data frame team_season_stats from MRegularSeasonDetailedResults

```{r}

rs <- data$MRegularSeasonDetailedResults

team_games <- bind_rows(
  rs %>%
    transmute(
      Season, DayNum,
      TeamID = WTeamID, OpponentID = LTeamID,
      Win = 1, NumOT,
      Points = WScore, OppPoints = LScore,
      FGM = WFGM, FGA = WFGA, FGM3 = WFGM3, FGA3 = WFGA3,
      FTM = WFTM, FTA = WFTA,
      OR = WOR, DR = WDR,
      Ast = WAst, TO = WTO,
      Stl = WStl, Blk = WBlk,
      PF = WPF
    ),
  rs %>%
    transmute(
      Season, DayNum,
      TeamID = LTeamID, OpponentID = WTeamID,
      Win = 0, NumOT,
      Points = LScore, OppPoints = WScore,
      FGM = LFGM, FGA = LFGA, FGM3 = LFGM3, FGA3 = LFGA3,
      FTM = LFTM, FTA = LFTA,
      OR = LOR, DR = LDR,
      Ast = LAst, TO = LTO,
      Stl = LStl, Blk = LBlk,
      PF = LPF
    )
)

team_season_stats <- team_games %>%
  group_by(Season, TeamID) %>%
  summarise(
    Games = n(),
    WinPct = mean(Win),
    AvgPoints = median(Points),
    AvgOppPoints = median(OppPoints),
    FG2_Pct = sum(FGM - FGM3) / sum(FGA - FGA3),
    FG3_Pct = sum(FGM3) / sum(FGA3),
    FT_Pct = sum(FTM) / sum(FTA),
    ThreeP_Rate = sum(FGA3) / sum(FGA),
    OR_pg = median(OR),
    DR_pg = median(DR),
    Ast_pg = median(Ast),
    TO_pg = median(TO),
    Stl_pg = median(Stl),
    Blk_pg = median(Blk),
    PF_pg = median(PF),
    AvgOT = mean(NumOT),
    .groups = "drop"
  )

# Filter to post-2002
team_season_stats <- team_season_stats %>%
  filter(Season >= 2003)
```

### Creating SOV and CareerTourneyWins

```{r}

# Strength of victories -----------------------------------------------------
# First, get the total number of teams per season for percentile calculation
teams_per_season <- kenpom_final %>%
  group_by(Season) %>%
  summarise(n_teams = n(), .groups = "drop")

# Prepare opponent KenPom ranks and join team count
opp_kenpom <- kenpom_final %>%
  select(Season, TeamID, Opp_RankAdjEM = RankAdjEM) %>%
  left_join(teams_per_season, by = "Season")

# Compute opponent percentile: (n_teams - rank + 1) / n_teams
opp_kenpom <- opp_kenpom %>%
  mutate(Opp_Percentile = (n_teams - Opp_RankAdjEM + 1) / n_teams) %>%
  select(Season, TeamID, Opp_Percentile)

# Join opponent percentiles to each game
team_games_with_opp <- team_games %>%
  left_join(opp_kenpom, by = c("Season", "OpponentID" = "TeamID"))

# Compute scaled SOV: average percentile of opponents that the team defeated
sov_scaled <- team_games_with_opp %>%
  filter(Win == 1) %>%
  group_by(Season, TeamID) %>%
  summarise(
    SOV_Scaled = mean(Opp_Percentile, na.rm = TRUE),
    .groups = "drop"
  )

# coach feature -----------------------------------------------------------
# Get all coach data from MTeamCoaches
all_coaches <- data$MTeamCoaches %>%
  select(Season, TeamID, CoachName) %>%
  distinct() %>%
  arrange(CoachName, Season)

# Get tournament wins for each coach-season
tourney_wins <- data$MNCAATourneyDetailedResults %>%
  # Get primary coach for the winning team
  inner_join(
    data$MTeamCoaches %>%
      mutate(DaysCoached = LastDayNum - FirstDayNum + 1) %>%
      group_by(Season, TeamID) %>%
      arrange(desc(DaysCoached)) %>%
      dplyr::slice(1) %>%
      ungroup() %>%
      select(Season, TeamID, PrimaryCoach = CoachName),
    by = c("Season", "WTeamID" = "TeamID")
  ) %>%
  # Count wins per primary coach per season
  group_by(PrimaryCoach, Season) %>%
  summarise(SeasonWins = n(), .groups = "drop") %>%
  rename(CoachName = PrimaryCoach)

# Calculate coach career statistics (using ALL historical data)
coach_career_stats <- all_coaches %>%
  select(CoachName, Season) %>%
  distinct() %>%
  left_join(tourney_wins, by = c("CoachName", "Season")) %>%
  mutate(SeasonWins = coalesce(SeasonWins, 0)) %>%
  arrange(CoachName, Season) %>%
  group_by(CoachName) %>%
  mutate(
    # Career wins BEFORE current season
    CareerTourneyWins = lag(cumsum(SeasonWins), default = 0),
    # Years coached BEFORE current season
    YearsNCAA = row_number() - 1,
    .groups = "drop"
  ) %>%
  select(CoachName, Season, YearsNCAA, CareerTourneyWins)

# Get primary coach for each team-season (for seasons >= 2003 to match team_season_stats)
primary_coaches <- data$MTeamCoaches %>%
  filter(Season >= 2003) %>%
  mutate(DaysCoached = LastDayNum - FirstDayNum + 1) %>%
  group_by(Season, TeamID) %>%
  arrange(desc(DaysCoached)) %>%
  dplyr::slice(1) %>%
  ungroup() %>%
  select(Season, TeamID, CoachName)


```

### Left joining strength of victory, coach historical success, and Kenpom metrics to team season stats

```{r}
team_season_stats <- team_season_stats %>%
  left_join(
    kenpom_final %>% 
      select(Season, TeamID, AdjEM, AdjOE, AdjDE, AdjTempo, RankAdjEM) %>%
      mutate(TeamID = as.integer(TeamID)),
    by = c("Season", "TeamID")
  )

team_season_stats <- team_season_stats %>%
  # Remove any existing coach columns if they exist
  select(-any_of(c("CoachName", "YearsNCAA", "CareerTourneyWins"))) %>%
  # Join with primary coaches
  left_join(primary_coaches, by = c("Season", "TeamID")) %>%
  # Join with career stats
  left_join(coach_career_stats, by = c("CoachName", "Season")) %>%
  # Fill NA values with 0
  mutate(
    YearsNCAA = coalesce(YearsNCAA, 0),
    CareerTourneyWins = coalesce(CareerTourneyWins, 0)
  )

team_season_stats <- team_season_stats %>%
  left_join(sov_scaled, by = c("Season", "TeamID")) %>%
  rename(SOV = SOV_Scaled)
```

Kenpom efficiency metrics had to be cleaned to be able to left join to team_season_stats (Appendix A)

## Summary Statistics

```{r summary-statistics}
skim(team_season_stats)
```

# Empirical Strategy / Methods

## Machine learning models

### Linear regression

For linear regression, we use the first two principle components resulting from the principal component analysis of the predictor variables.

$$Probability_{i} = \beta_0 + \beta_1 PC_1 X_{i} + \beta_2 PC_2 X_{i} + \varepsilon_{i}$$

where $Probability_{i}$ is the outcome of win probability for unit $i$, $PC_1X_{it}$ is the first principal component of the predictor variables, $PC_2X_{it}$ is the second explanatory of the predictor variables, and $\varepsilon_{i}$ is the error term. We then cap the maximum probability at 100% and the minimum at 0%.

### Logistic regression

$$p(Y=1 \mid X_1, X_2)
=
\frac{1}{1 + \exp\left(-\left(\beta_0 + \sum_{j=1}^p \beta_j (X_{j,1} - X_{j,2})\right)\right)}$$

$X_{j,1}$ and $X_{j,2}$ are team-level stats for team 1 and team 2.

### Random Forest

### Elastic Net

### Training the models

For each year the model is to predict, the model gets retrained on all previous years.

## Public Picks (comparison model)

We calculate a raw strength value based on the frequency a team was picked to each round of the tournament using the following formula:

$$Strength = r_1 + 2r_2 + 4r_3 + 8r_4 + 16r_5 + 32r_6$$

We normalize the strength to each season by dividing each strength value by the highest strength value for the given year. $$Strength_{\mathrm{norm}} = \frac{Strength}{\max(Strength)}$$

We convert the normalized strength value into an ELO metric

$$\mathrm{ELO} = 1500 + 400\ln\left(Strength_{\mathrm{norm}} + 10^{-6}\right)$$

We finally compare the ELO of two teams to determine the probability a given team wins

$$P(\text{A beats B}) = \frac{1}{1 + 10^{\frac{ELO_B - ELO_A}{500}}}$$

Where $A$ refers to team A, and $B$ refers to team B.

## Evaluation methodology

### Brier Score:

$$\text{Brier Score} = \frac{1}{N}\sum_{i=1}^{N}(p_i - o_i)^2$$

- $p_i$ = predicted probability
- $o_i$ = observed outcome (0 for a loss, 1 for a win)
- \$N\$ = number of predictions

### Log-Loss:

$$\text{Log Loss} = -\frac{1}{N}\sum_{i=1}^{N}\left[o_i\log(p_i) + (1-o_i)\log(1-p_i)\right]$$

- $p_i$ = predicted probability
- $o_i$ = observed outcome (0 for a loss, 1 for a win)
- $N$ = number of predictions

### Simulated Brackets

To evaluate the models on simulated brackets, first we run simulations with the following method:

$$Y_{g,b} \sim \operatorname{Bernoulli}(p_{g,b}), \qquad g=1,\dots,63,\ b=1,\dots,100$$

$Y_{g,b}$ = the simulated outcome of game $g$ in bracket simulation $b$

$Bernoulli(p_{g,b})$ = a random draw with probability $p_{g,b}​$

$$B_b = \{Y_{1,b},Y_{2,b},\dots,Y_{63,b}\}$$

$B_b$ = one complete simulated tournament bracket with all random draws chosen for all 63 games

These simulation are run 100 times

#### Scoring brackets

The brackets are scored on the following formula:

$$S({B}) = \sum_{r=1}^{6} w_r C_r$$

$$(w_1,w_2,w_3,w_4,w_5,w_6) = (1,1,2,3,5,8)$$

$S(B)$ = Total score for bracket $B$

$C_r$ = Total number of correct picks for round $r$

$w_r$ = point value for round $r$, ranging from 1 point for the first round up to 8 points for the championship game

This gives a score for each bracket each model generates, which we can use to compare to other models

# Results

Present your main empirical findings using tables, figures, and written interpretation.

```{r results-table}
# Replace this example with your own table or model output.
# modelsummary(list(model1, model2))
```

```{r results-figure}
# Replace this example with your own figure.
# ggplot(your_data, aes(x = x_variable, y = y_variable)) +
#   geom_point() +
#   labs(x = "X Variable", y = "Y Variable", title = "Main Result")
```

# Discussion and Implications

Interpret your results. Explain what they imply for economics, business, policy, or society.

Discuss limitations and clarify what your analysis can and cannot show.

# Conclusion

Summarize your research question, data, method, main findings, and key takeaway.

Briefly suggest possible future extensions.

# References

Shayer, E., & Powers, S. (2017). *mRchmadness: Numerical tools for filling out an NCAA basketball tournament bracket* (Version 1.0.0) \[R package\]. CRAN. https://CRAN.R-project.org/package=mRchmadness. Accessed 4/26/2026.

# Appendix

## Appendix A

### Kenpom Cleaning

```{r}
# --- 1. LOAD KENPOM DATA WITH ALL STATS ---
kenpom1 <- data$INT_KenPom_Summary_Pre_Tourny %>%
  select(Season, TeamName, AdjEM, RankAdjEM, AdjOE, AdjDE, AdjTempo) %>%
  mutate(TeamName = tolower(trimws(TeamName)))

#---------- 2026 years kenpom
kenpom2 <- data$INT_KenPom_Summary_2026 %>% 
  select(Season, TeamName, AdjEM, RankAdjEM, AdjOE, AdjDE, AdjTempo) %>%
  filter(Season==2026) %>% 
  mutate(TeamName = tolower(trimws(TeamName)))

kenpom <- bind_rows(kenpom1,kenpom2)

# Load spellings and team info
spellings <- data$MTeamSpellings %>%
  mutate(TeamNameSpelling = tolower(trimws(
    iconv(TeamNameSpelling, from = "latin1", to = "UTF-8", sub = " ")
  ))) %>%
  distinct(TeamNameSpelling, .keep_all = TRUE)

# --- 2. COMPREHENSIVE MANUAL MAPPINGS ---
# These handle all known name variations and historical name changes
manual_mappings <- data.frame(
  TeamName = c(
    # Louisiana Lafayette variations
    "louisiana lafayette", "louisiana-lafayette", "ull",
    # Illinois Chicago variations
    "illinois chicago", "illinois-chicago", "uic",
    # Texas A&M Corpus Christi
    "texas a&m corpus chris", "texas a&m-corpus christi", "a&m-corpus chris", "texas a&m-cc", "corpus christi",
    # Mississippi Valley State
    "mississippi valley st.", "mississippi valley state", "miss valley st.",
    # Arkansas Pine Bluff
    "arkansas pine bluff", "ark pine bluff", "uapb",
    # Arkansas Little Rock
    "arkansas little rock", "ark little rock", "ualr", "little rock",
    # Cal State Bakersfield
    "cal st. bakersfield", "cal state bakersfield", "csu bakersfield", "cs bakersfield", "bakersfield",
    # Southeast Missouri State
    "southeast missouri st.", "southeast missouri state", "se missouri st", "se missouri state", "semo",
    # Texas State
    "texas state", "texas st.", "southwest texas st.", "southwest texas state",
    # UT Rio Grande Valley
    "ut rio grande valley", "texas-rio grande valley", "utrgv", "texas pan american", "texas-pan american",
    # IUPUI / IU Indy
    "iupui", "iu indy", "indianapolis",
    # LIU
    "liu", "long island", "long island university", "liu brooklyn",
    # Queens
    "queens", "queens nc", "queens university",
    # Saint Francis (NY)
    "st. francis ny", "st francis ny", "saint francis ny", "st. francis (ny)", "saint francis (ny)",
    # Saint Francis (PA)
    "st. francis pa", "st francis pa", "saint francis pa", "st. francis (pa)", "saint francis (pa)",
    # Green Bay
    "green bay", "wisconsin green bay", "wisconsin-green bay",
    # Milwaukee
    "milwaukee", "wisconsin milwaukee", "wisconsin-milwaukee",
    # Utah Tech / Dixie State
    "utah tech", "dixie st.", "dixie state",
    # Texas A&M Commerce
    "texas a&m commerce", "tx a&m commerce", "east texas a&m",
    # Missouri State
    "missouri state", "missouri st.", "southwest missouri st.",
    # McNeese
    "mcneese", "mcneese st.", "mcneese state",
    # Nicholls
    "nicholls", "nicholls st.", "nicholls state",
    # Troy
    "troy", "troy st.",
    # Utah Valley
    "utah valley", "utah valley st.", "utah valley state",
    # Detroit Mercy
    "detroit mercy", "detroit", "detroit-mercy",
    # SIU Edwardsville
    "siue", "siu edwardsville", "southern illinois edwardsville",
    # UMKC / Kansas City
    "umkc", "kansas city", "missouri-kansas city",
    # Purdue Fort Wayne
    "purdue fort wayne", "ipfw", "fort wayne", "pfw",
    # Charleston
    "charleston", "college of charleston",
    # CSUN
    "csun", "cal st. northridge", "cal state northridge",
    # Middle Tennessee
    "middle tennessee", "middle tennessee st.", "mtsu",
    # Additional tournament-relevant teams
    "louisiana monroe", "st francis (pa)", "bethune-cookman", 
    "maryland - e. shore", "tarleton st", "tarleton state",
    "st. thomas", "tarleton st.", "winston salem st.", "tennessee martin"
  ),
  TeamID = c(
    rep(1418, 3), rep(1227, 3), rep(1394, 5), rep(1290, 3),
    rep(1115, 3), rep(1114, 4), rep(1167, 5), rep(1369, 5),
    rep(1402, 4), rep(1410, 5), rep(1237, 3), rep(1254, 4),
    rep(1474, 3), rep(1383, 5), rep(1382, 5),
    rep(1453, 3), rep(1454, 3), rep(1469, 3), rep(1477, 3),
    rep(1283, 3), rep(1270, 3), rep(1311, 3), rep(1407, 2),
    rep(1430, 3), rep(1178, 3), rep(1188, 3), rep(1282, 3),
    rep(1236, 4), rep(1158, 2), rep(1169, 3), rep(1292, 3),
    # Additional IDs
    1398, 1384, 1126, 1271, 1470, 1470, 1384, 1468, 1475, 1404
  ),
  stringsAsFactors = FALSE
) %>%
  distinct(TeamName, .keep_all = TRUE)

# --- 3. EXACT MATCHES ---
exact <- kenpom %>%
  left_join(spellings, by = c("TeamName" = "TeamNameSpelling")) %>%
  filter(!is.na(TeamID)) %>%
  mutate(method = "exact")

# --- 4. MANUAL MATCHES (apply first) ---
manual <- kenpom %>%
  filter(TeamName %in% manual_mappings$TeamName) %>%
  left_join(manual_mappings, by = "TeamName") %>%
  mutate(method = "manual")

# --- 5. FUZZY MATCH FOR REMAINING ---
already_matched <- unique(c(exact$TeamName, manual$TeamName))
fuzzy_candidates <- kenpom %>%
  filter(!TeamName %in% already_matched) %>%
  distinct(TeamName) %>%
  pull(TeamName)

fuzzy_list <- list()
for (team in fuzzy_candidates) {
  dists <- stringdist(team, spellings$TeamNameSpelling, method = "jw")
  best <- which.min(dists)
  if (dists[best] <= 0.2) {
    # Get the matched spelling and TeamID
    matched_spelling <- spellings$TeamNameSpelling[best]
    matched_id <- spellings$TeamID[best]
    
    # Retrieve the full KenPom row for that team (Season, AdjEM, RankAdjEM, AdjOE, AdjDE, AdjTempo)
    full_row <- kenpom %>% filter(TeamName == team)
    if (nrow(full_row) > 0) {
      fuzzy_list[[team]] <- full_row %>%
        mutate(
          TeamID = matched_id,
          matched_spelling = matched_spelling,
          dist = dists[best],
          method = "fuzzy"
        )
    }
  }
}
fuzzy <- bind_rows(fuzzy_list)

# --- 6. COMBINE ALL MATCHES ---
# Join back to full KenPom data (keeping all stats including RankAdjEM)
kenpom_final <- bind_rows(
  exact %>% select(Season, TeamName, AdjEM, RankAdjEM, AdjOE, AdjDE, AdjTempo, TeamID, method),
  manual %>% select(Season, TeamName, AdjEM, RankAdjEM, AdjOE, AdjDE, AdjTempo, TeamID, method),
  fuzzy %>% select(Season, TeamName, AdjEM, RankAdjEM, AdjOE, AdjDE, AdjTempo, TeamID, method)
) %>%
  distinct(Season, TeamID, .keep_all = TRUE) %>%   # one row per team-season
  arrange(Season, TeamID)

# --- 8. FINAL VERIFICATION ---
cat("\n", strrep("=", 60), "\n")
cat("FINAL KENPOM DATASET\n")
cat(strrep("=", 60), "\n")
cat("Rows:", nrow(kenpom_final), "\n")
cat("Unique teams:", n_distinct(kenpom_final$TeamID), "\n")
cat("Seasons:", min(kenpom_final$Season, na.rm=TRUE), "-", max(kenpom_final$Season, na.rm=TRUE), "\n")

# Verify tournament teams (2003+)
tourney_teams <- data$MNCAATourneySeeds %>%
  filter(Season >= 2003) %>%
  distinct(TeamID) %>%
  pull(TeamID)

missing <- setdiff(tourney_teams, kenpom_final$TeamID)
if (length(missing) == 0) {
  cat("\n✅ TOURNAMENT TEAMS (2003+): ALL", length(tourney_teams), "PRESENT\n")
} else {
  cat("\n⚠️ Missing tournament teams:", length(missing), "\n")
  cat("Missing IDs:", paste(missing, collapse = ", "), "\n")
}
```
