---
title: Data Analytics Capstone Project
subtitle: "NCAA March Madness Prediction"
author:
  - name: Aaron Zalen 
    affiliation: SUNY Geneseo
    
date: last-modified

execute: 
  echo: true
  eval: true
  warning: false
  message: false
  fig-width: 10
  fig-height: 6
  dir: project

format:
  html:
    code-fold: true
    toc: true
    toc-title: "Contents"
    toc-expand: 3
    number-sections: true
    code-tools: true
    highlight-style: atom-one
---

```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(rmarkdown)
library(DT)
library(dplyr)
library(readr)
library(purrr)
library(tibble)
library(ranger)  
library(stringr)
library(tidyr)
library(glmnet)
library(dplyr)
library(stringdist)
library(readr)
library(purrr)
library(tibble)
# library(randomForest)
library(stringr)
library(tidyr)
library(glmnet)
library(tidyverse)

# Feel free to revise this ggplot theme_set
theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = ggplot2::margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 90,
                               size = rel(1.33),
                               margin = ggplot2::margin(0,10,0,0))
                )
          )


# Feel free to use below global options for color-blind-friendly scales
# scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
# scale_fill_discrete <- function(...) scale_fill_viridis_d(...)
```

## This page is currently under construction

# Introduction

Describe the real-world motivation for your project. Clearly state your research question and explain why it matters.

Briefly preview your data, empirical approach, main findings, and contribution.

March Madness is an annual tournament for division 1 college basketball. I will be training prediction models to assign a probability to teams winning in a match up. One of the prediction models is derived from human-selected bracket data sourced from ESPN to compare the results of the machine learning models against.

The research question is "How do NCAA March Madness prediction models perform relative to human-selected brackets?"

I will be evaluating each model on Brier-score, Log-loss, and Scored Simulated Brackets

The goal is to find a way to fill out brackets which consistently outperforms both people and the other prediction models. If a model is able to consistently outperform people at filling out brackets, anyone who relies on that model will have an advantage.

# Background and Related Literature

The structure of March Madness is a 64 team, single elimination tournament.

Millions of people fill out brackets each year, attempting to predict which teams will move on to each round. Billions of dollars each year are gambled on the tournament.

# Data

## Data Source

Regular season performance data and bracket structure data: Kaggle

Advanced efficiency metrics: Kenpom

Public picks data: mRchmadness package, which scrapes data from ESPN

## Unit of Observation

Each observation represents team-level pre-tournament aggregate data

## Key Variables

Describe your outcome variable, main explanatory variables, and important control variables.

The outcome variable we are predicting for is win probability.

WinPct= Win percentage

AvgPoints= Average points

AvgOppPoints= Average Opponent Points

FG2_Pct= two-point percentage

FG3_Pct= three-point percentage

FT_Pct= free-throw percentage

OR_pg= Offensive rebounds per game

DR_pg= Defensive rebounds per game

TO_pg= Turnovers per game

CareerTourneyWins= Coach's historical success in previous tournaments, measured in number of NCAA tournament games won.

AdjEM= Adjusted efficiency margin, the number of points a team is expected to outscore the average D-1 team over 100 possessions

SOV= Strength of victory, average AdjEM percentile of opponents that a team defeated

AdjOE= Adjusted offensive efficiency, sourced from kenpom, points scored per 100 possessions and adjusted for opponent.

AdjDE= Adjusted defensive efficiency, points allowed per 100 possessions and adjusted for opponent.

RankAdjEM= Ranked adjusted efficiency margin, the number ranking of each team's Adjusted efficiency margin each season

AdjTempo= Adjusted tempo, measured in possessions per 40 minutes and adjusted for opponent.

## Data Cleaning

Explain major cleaning steps, sample restrictions, missing values, and variable construction.

loading data into a directory

```{r}
data_dir <- "csvs"

files <- list.files(
  data_dir,
  pattern = "\\.csv$",
  full.names = TRUE
)

data <- files %>%
  set_names(tools::file_path_sans_ext(basename(.))) %>%
  map(~ tryCatch(
    read_csv(.x, show_col_types = FALSE),
    error = function(e) {
      message("Error reading: ", .x)
      NULL
    }
  ))
```

```{r}

rs <- data$MRegularSeasonDetailedResults

team_games <- bind_rows(
  rs %>%
    transmute(
      Season, DayNum,
      TeamID = WTeamID, OpponentID = LTeamID,
      Win = 1, NumOT,
      Points = WScore, OppPoints = LScore,
      FGM = WFGM, FGA = WFGA, FGM3 = WFGM3, FGA3 = WFGA3,
      FTM = WFTM, FTA = WFTA,
      OR = WOR, DR = WDR,
      Ast = WAst, TO = WTO,
      Stl = WStl, Blk = WBlk,
      PF = WPF
    ),
  rs %>%
    transmute(
      Season, DayNum,
      TeamID = LTeamID, OpponentID = WTeamID,
      Win = 0, NumOT,
      Points = LScore, OppPoints = WScore,
      FGM = LFGM, FGA = LFGA, FGM3 = LFGM3, FGA3 = LFGA3,
      FTM = LFTM, FTA = LFTA,
      OR = LOR, DR = LDR,
      Ast = LAst, TO = LTO,
      Stl = LStl, Blk = LBlk,
      PF = LPF
    )
)

team_season_stats <- team_games %>%
  group_by(Season, TeamID) %>%
  summarise(
    Games = n(),
    WinPct = mean(Win),
    AvgPoints = median(Points),
    AvgOppPoints = median(OppPoints),
    FG2_Pct = sum(FGM - FGM3) / sum(FGA - FGA3),
    FG3_Pct = sum(FGM3) / sum(FGA3),
    FT_Pct = sum(FTM) / sum(FTA),
    ThreeP_Rate = sum(FGA3) / sum(FGA),
    OR_pg = median(OR),
    DR_pg = median(DR),
    Ast_pg = median(Ast),
    TO_pg = median(TO),
    Stl_pg = median(Stl),
    Blk_pg = median(Blk),
    PF_pg = median(PF),
    AvgOT = mean(NumOT),
    .groups = "drop"
  )


```

```{r}
kenpom_final <- data$kenpom_tournament_ready
# Filter to post-2002
team_season_stats <- team_season_stats %>%
  filter(Season >= 2003)
```

```{r}

# Strength of victories -----------------------------------------------------
# First, get the total number of teams per season for percentile calculation
teams_per_season <- kenpom_final %>%
  group_by(Season) %>%
  summarise(n_teams = n(), .groups = "drop")

# Prepare opponent KenPom ranks and join team count
opp_kenpom <- kenpom_final %>%
  select(Season, TeamID, Opp_RankAdjEM = RankAdjEM) %>%
  left_join(teams_per_season, by = "Season")

# Compute opponent percentile: (n_teams - rank + 1) / n_teams
opp_kenpom <- opp_kenpom %>%
  mutate(Opp_Percentile = (n_teams - Opp_RankAdjEM + 1) / n_teams) %>%
  select(Season, TeamID, Opp_Percentile)

# Join opponent percentiles to each game
team_games_with_opp <- team_games %>%
  left_join(opp_kenpom, by = c("Season", "OpponentID" = "TeamID"))

# Compute scaled SOV: average percentile of opponents that the team defeated
sov_scaled <- team_games_with_opp %>%
  filter(Win == 1) %>%
  group_by(Season, TeamID) %>%
  summarise(
    SOV_Scaled = mean(Opp_Percentile, na.rm = TRUE),
    .groups = "drop"
  )

# coach feature -----------------------------------------------------------
# Get all coach data from MTeamCoaches
all_coaches <- data$MTeamCoaches %>%
  select(Season, TeamID, CoachName) %>%
  distinct() %>%
  arrange(CoachName, Season)

# Get tournament wins for each coach-season
tourney_wins <- data$MNCAATourneyDetailedResults %>%
  # Get primary coach for the winning team
  inner_join(
    data$MTeamCoaches %>%
      mutate(DaysCoached = LastDayNum - FirstDayNum + 1) %>%
      group_by(Season, TeamID) %>%
      arrange(desc(DaysCoached)) %>%
      dplyr::slice(1) %>%
      ungroup() %>%
      select(Season, TeamID, PrimaryCoach = CoachName),
    by = c("Season", "WTeamID" = "TeamID")
  ) %>%
  # Count wins per primary coach per season
  group_by(PrimaryCoach, Season) %>%
  summarise(SeasonWins = n(), .groups = "drop") %>%
  rename(CoachName = PrimaryCoach)

# Calculate coach career statistics (using ALL historical data)
coach_career_stats <- all_coaches %>%
  select(CoachName, Season) %>%
  distinct() %>%
  left_join(tourney_wins, by = c("CoachName", "Season")) %>%
  mutate(SeasonWins = coalesce(SeasonWins, 0)) %>%
  arrange(CoachName, Season) %>%
  group_by(CoachName) %>%
  mutate(
    # Career wins BEFORE current season
    CareerTourneyWins = lag(cumsum(SeasonWins), default = 0),
    # Years coached BEFORE current season
    YearsNCAA = row_number() - 1,
    .groups = "drop"
  ) %>%
  select(CoachName, Season, YearsNCAA, CareerTourneyWins)

# Get primary coach for each team-season (for seasons >= 2003 to match team_season_stats)
primary_coaches <- data$MTeamCoaches %>%
  filter(Season >= 2003) %>%
  mutate(DaysCoached = LastDayNum - FirstDayNum + 1) %>%
  group_by(Season, TeamID) %>%
  arrange(desc(DaysCoached)) %>%
  dplyr::slice(1) %>%
  ungroup() %>%
  select(Season, TeamID, CoachName)


```

Adding strength of victory, coach historical success, and Kenpom metrics to team season stats

```{r}
team_season_stats <- team_season_stats %>%
  left_join(
    kenpom_final %>% 
      select(Season, TeamID, AdjEM, AdjOE, AdjDE, AdjTempo, RankAdjEM) %>%
      mutate(TeamID = as.integer(TeamID)),
    by = c("Season", "TeamID")
  )

team_season_stats <- team_season_stats %>%
  # Remove any existing coach columns if they exist
  select(-any_of(c("CoachName", "YearsNCAA", "CareerTourneyWins"))) %>%
  # Join with primary coaches
  left_join(primary_coaches, by = c("Season", "TeamID")) %>%
  # Join with career stats
  left_join(coach_career_stats, by = c("CoachName", "Season")) %>%
  # Fill NA values with 0
  mutate(
    YearsNCAA = coalesce(YearsNCAA, 0),
    CareerTourneyWins = coalesce(CareerTourneyWins, 0)
  )

team_season_stats <- team_season_stats %>%
  left_join(sov_scaled, by = c("Season", "TeamID")) %>%
  rename(SOV = SOV_Scaled)
```

Kenpom efficiency metrics had to be cleaned to be able to left join to team_season_stats (Appendix A)

## Summary Statistics

```{r summary-statistics}
# Replace this example with your own data and summary statistics.
# skim(your_data)
```

# Empirical Strategy / Methods

Describe the method used to answer your research question. This may include regression, classification, clustering, forecasting, text analysis, or another data analytics method.

Linear Regression Using PC1 and PC2:

$$Probability_{i} = \beta_0 + \beta_1 PC_1 X_{i} + \beta_2 PC_2 X_{i} + \varepsilon_{i}$$

where $Probability_{i}$ is the outcome of win probability for unit $i$, $PC_1X_{it}$ is the first explanatory variable, $PC_2X_{it}$ is the second explanatory variable, and $\varepsilon_{i}$ is the error term. We then cap the maximum probability at 100% and the minimum at 0%.

The methodology to create a public consensus-based model is as follows:

We calculate a raw strength value based on the frequency a team was picked to each round of the tournament using the following formula: $$Strength = r_1 + 2r_2 + 4r_3 + 8r_4 + 16r_5 + 32r_6$$ We normalize the strength to each season by dividing each strength value by the highest strength value for the given year. $$Strength_{\mathrm{norm}} = \frac{Strength}{\max(Strength)}$$

We convert the normalized strength value into an ELO metric

$$\mathrm{ELO} = 1500 + 400\ln\left(Strength_{\mathrm{norm}} + 10^{-6}\right)$$

We finally compare the ELO of two teams to determine the probability a given team wins

$$P(\text{A beats B}) = \frac{1}{1 + 10^{\frac{ELO_B - ELO_A}{500}}}$$

Where $A$ refers to team A, and $B$ refers to team B.

Logistic regression: $$p(Y=1 \mid X_1, X_2)
=
\frac{1}{1 + \exp\left(-\left(\beta_0 + \sum_{j=1}^p \beta_j (X_{j,1} - X_{j,2})\right)\right)}$$

$X_{j,1}$ and   $X_{j,2}$ are team-level stats for team 1 and team 2.

If your project uses machine learning or prediction, explain your train/test split, cross-validation, model tuning, and performance metrics.

# Results

Present your main empirical findings using tables, figures, and written interpretation.

```{r results-table}
# Replace this example with your own table or model output.
# modelsummary(list(model1, model2))
```

```{r results-figure}
# Replace this example with your own figure.
# ggplot(your_data, aes(x = x_variable, y = y_variable)) +
#   geom_point() +
#   labs(x = "X Variable", y = "Y Variable", title = "Main Result")
```

# Discussion and Implications

Interpret your results. Explain what they imply for economics, business, policy, or society.

Discuss limitations and clarify what your analysis can and cannot show.

# Conclusion

Summarize your research question, data, method, main findings, and key takeaway.

Briefly suggest possible future extensions.

# References

List all sources cited in your project, including data sources, academic articles, reports, websites, packages, and tools.

For online sources, include the full URL and date/time accessed.

# Appendix

Include additional tables, figures, robustness checks, code explanations, or variable definitions if needed.
