---
title: Data Analytics Capstone Title
subtitle: "My Project"
author:
  - name: YOUR NAME 
    affiliation: SUNY Geneseo
    
date: last-modified

execute: 
  echo: true
  eval: true
  warning: false
  message: false
  fig-width: 10
  fig-height: 6

format:
  html:
    code-fold: true
    toc: true
    toc-title: "Contents"
    toc-expand: 3
    number-sections: true
    code-tools: true
    highlight-style: atom-one
---

```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(rmarkdown)
library(DT)
library(dplyr)
library(readr)
library(purrr)
library(tibble)
library(ranger)  
library(stringr)
library(tidyr)
library(glmnet)
library(dplyr)
library(stringdist)
library(readr)
library(purrr)
library(tibble)
# library(randomForest)
library(stringr)
library(tidyr)
library(glmnet)

# Feel free to revise this ggplot theme_set
theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = ggplot2::margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 90,
                               size = rel(1.33),
                               margin = ggplot2::margin(0,10,0,0))
                )
          )


# Feel free to use below global options for color-blind-friendly scales
# scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
# scale_fill_discrete <- function(...) scale_fill_viridis_d(...)
```

## This page is currently under construction

# Introduction

Describe the real-world motivation for your project. Clearly state your research question and explain why it matters.

Briefly preview your data, empirical approach, main findings, and contribution.

March Madness is an annual tournament for division 1 college basketball. I will be training prediction models to assign a probability to teams winning in a match up. One of the prediction models is derived from human-selected bracket data sourced from ESPN.

The research question is "How do NCAA March Madness prediction models perform relative to human-selected brackets?"

The goal is to find a way to fill out brackets which consistently outperforms both people and the other prediction models. If a model is able to consistently outperform people at filling out brackets, anyone who relies on that model will have an advantage.

# Background and Related Literature

-64 teams compete for the title

-Single elimination tournament.

-Millions of people fill out brackets

-Billions of dollars each year are gambled on the tournament

# Data

## Data Source

Describe the data source clearly. Include the source name, link, and access date in the References section.

## Unit of Observation

Explain what each row in your dataset represents.

## Key Variables

Describe your outcome variable, main explanatory variables, and important control variables.

The outcome variable is win probability.

## Data Cleaning

Explain major cleaning steps, sample restrictions, missing values, and variable construction.

Kenpom had to be cleaned to be able to left join to team_season_stats.

## Summary Statistics

```{r summary-statistics}
# Replace this example with your own data and summary statistics.
# skim(your_data)
```

# Empirical Strategy / Methods

Describe the method used to answer your research question. This may include regression, classification, clustering, forecasting, text analysis, or another data analytics method.

Linear Regression Using PC1 and PC2:

$$Probability_{i} = \beta_0 + \beta_1 PC_1 X_{i} + \beta_2 PC_2 X_{i} + \varepsilon_{i}$$

where $Probability_{i}$ is the outcome of win probability for unit $i$, $PC_1X_{it}$ is the first explanatory variable, $PC_2X_{it}$ is the second explanatory variable, and $\varepsilon_{i}$ is the error term. We then cap the maximum probability at 100% and the minimum at 0%.

If your project uses machine learning or prediction, explain your train/test split, cross-validation, model tuning, and performance metrics.

# Results

Present your main empirical findings using tables, figures, and written interpretation.

```{r results-table}
# Replace this example with your own table or model output.
# modelsummary(list(model1, model2))
```

```{r results-figure}
# Replace this example with your own figure.
# ggplot(your_data, aes(x = x_variable, y = y_variable)) +
#   geom_point() +
#   labs(x = "X Variable", y = "Y Variable", title = "Main Result")
```

# Discussion and Implications

Interpret your results. Explain what they imply for economics, business, policy, or society.

Discuss limitations and clarify what your analysis can and cannot show.

# Conclusion

Summarize your research question, data, method, main findings, and key takeaway.

Briefly suggest possible future extensions.

# References

List all sources cited in your project, including data sources, academic articles, reports, websites, packages, and tools.

For online sources, include the full URL and date/time accessed.

# Appendix

Include additional tables, figures, robustness checks, code explanations, or variable definitions if needed.
